TrenchR: an R package for transparent environmental and ecological biophysics

### 2020-11-10 Manuscript_full.Rmd
Update ms

## Background
Many responses of species and ecosystems to climate change are inconsistent with predictive most current models [@maguire2015modeling]. Some of these predictive shortcomings stem from omitting spatial and temporal environmental variation and how it interacts with organismal phenotypes (e.g., size, coloration) [@nadeau2017coarse; @potter2013microclimatic]. Many analyses assume that organismal body temperatures are equal to air temperate at weather station height (usually 2m). However, air temperatures near the ground where most organisms reside can be considerably warmer and absorption of solar radiation can raise body temperatures well above air temperatures, resulting in potential discrepancies of tens of degrees C [Kearney et al., 2009; Sinclair et al., 2016; Sunday et al., 2014].  Acute thermal stress events are often more relevant to the physiology and energy balance of organisms than daily, monthly, or annually aggregated environmental conditions. Yet, daily and seasonal environmental variation is often neglected [@dillon2016introduction; @dillon2016life]. We aim to help remedy these shortcomings by introducing the TrenchR package as a component of the TrEnCh project (https://trenchproject.github.io), which builds computational and visualization tools to Translate Environmental Change into organismal responses.

The mathematical tools for leveraging environmental data to predict organismal conditions have long lingered in books [@campbell2000introduction; @gates1980biophysical] and articles [@porter1969thermodynamic; @porter1983biophysical], but adoption of the tools has not kept pace with research on climate change responses. Adoptions of biophysical approaches are expanding with recognition of the importance of spatial and temporal environmental variation to determining climate change responses [@buckley2018leveraging; @kearney2017nichemapr].  Increasing dissemination of computational tools is facilitating adoption. The TrenchR package is intended to complement the NicheMapR package [@kearney2017nichemapr], which includes sophisticated microclimate and biophysical models. Many of the approaches are similar between the packages, but TrenchR focuses on simple functions to aid the understanding and transparency of biophysical approaches.  The TrenchR functions allow for modular use and easy adaptation to a variety of organismal and research questions.  Some energy budget calculation components of biophysical models are available in the ThermImage package (Tattersall, 2017). Although many of the TrenchR functions are general, some components of TrenchR are best suited for ectothermic animals. We have minimized inclusion of plant specific biophysical models since they are the focus of several R packages [tealeaves: (Muir, 2019); plantecophys [@duursma2015plantecophys]].

@@

## Examples
We illustrate the use of the TrenchR package by estimating an energy budget for a *Sceloporus* lizard on June 1, 2018, in Santa Fe, New Mexico, USA (35.69, -105.944, elevation: 2121m). The example is also incorporated in the *Using energy balances to estimate body temperatures* tutorial. We first estimate the energy budget with component functions and we then use an integrated biophysical model to estimate operative environmental temperatures. We start by generating environmental inputs.

@@

Let's assume the lizard is in an unshaded location where the daily air temperature varies from a minimum of 10°C to a maximum of 25°C, the soil surface temperature varies from a minimum of 15°C to a maximum of 30°C, and the windspeed is 0.5 m/s. We assume that atmospheric transmissivity $\tau=0.7$ and albedo $\rho=0.6$

@@

#set up input data as variables
lat=35.69; lon= -105.944; elev=2121 #latitude and longitude (degrees) and elevation (m)
Tmin=10; Tmax=25; Tmin_s=15; Tmax_s=30 #minimum and maximum of air and soil temperatures, respectively (C)
V=0.5 #wind speed (m/s)
#assumptions
tau=0.7; rho=0.6 #atmospheric transmissivity and albedo
Tb0=25 #initial assumption of body temperature (C)

@@

We can estimate hourly solar radiation as follows:
```{r}
psi_deg= sapply(1:24, FUN=zenith_angle, doy=doy, lat=lat, lon=lon) #estimate zenith angle (degrees)
psi_rad= degree_to_radian(psi_deg) #convert to radians
Srad= sapply(psi_rad, FUN=estimate_radiation, doy=doy, tau=tau, elev=elev, rho=rho) #estimate radiation
#Separate into direct, diffuse, and reflected solar radiation
Sdir= Srad[1,]
Sdif= Srad[2,]
Sref= Srad[3,]
```

We can then calculate hourly air and soil temperatures. We use the sine-exponential model for air temperature and the sine model for surface temperature:
```{r}
Ta= sapply(1:24, diurnal_temp_variation_sineexp, T_max=Tmax, T_min=Tmin, t_r=tr, t_s=ts, alpha=2.59, beta= 1.55, gamma=2.2)
Ts= sapply(1:24, diurnal_temp_variation_sine, T_max=Tmax_s, T_min=Tmin_s)
```

### Energy balance
We will be solving the following energy balance to estimate $T_e$:
$$Q_{net} = Q_{abs} - Q_{emit} + Q_{conv} + Q_{cond} + Q_{met} - Q_{evap},$$
We will estimate each term on the right side of the equation in turn. Estimating $Q_{abs}$requires the surface area exposed to radiation and the solar absorptivity of the animal surface ($a$ proportion). We model a 10g Sceloporus lizard with solar absorptivity $a=0.9$ [@gates1980biophysical]. We will initially assume $T_b=T_a$ to illustrate calculations.

```{r}
mass=10
a=0.9 #solar absorptivity

@@

We calculate the hourly radiation absorbed as follows:
```{r}
Qabs= rep(NA, 24)
for(hr in 1:24) Qabs[hr]= Qradiation_absorbed(a=a, A=A, psa_dir=psa[hr], psa_ref=1-psa[hr], S_dir=Sdir[hr], S_dif=Sdif[hr], a_s=rho)
```

We estimate thermal radiation $Q_{emit}$ for the lizard outdoors. We assume the surface emissivity of lizards, $epsilon_s=0.965$ [Bartlett & Gates 1967].

@@

Qemit= rep(NA, 24)
for(hr in 1:24) Qemit[hr]=Qemitted_thermal_radiation(epsilon=epsilon_s, A=A, psa_dir=psa[hr], psa_ref=1-psa[hr], T_b=Ta[hr]+273.15, T_g=Ts[hr]+273.15, T_a=Ta[hr]+273.15, enclosed=FALSE)
```

We next estimate convection $Q_{conv}$ and conduction $Q_{cond}$. We will estimate the lizard's heat transfer coefficient, $H_L$ using an empirical relationship for lizards (`heat_transfer_coefficient()`), but also illustrate a function estimating $H_L$ using a spherical approximation (`heat_transfer_coefficient_approximation()`) and a simplified approximation (`heat_transfer_coefficient_simple()`).

```{r}
#Use DRYAIR in NicheMapR to estimate the thermal conductivity of air and kinematic viscosity
#library(NicheMapR)

#estimate Barometric pressure (pascal)
ap= airpressure_from_elev(elev)*1000

DRYAIRout= DRYAIR(db=Ta, bp=ap, alt=elev)
K= DRYAIRout$thcond
nu= DRYAIRout$viskin #Kinematic viscosity (m2 s-1)

svl=0.006 #SVL for sceloporus

#We will use the average of K and nu across the day for simplicity and since there's not a substantial variation
K= mean(K)
nu=mean(nu)

#Estimate using an empirical relationship
H_L=heat_transfer_coefficient(V=V,D=svl,K= 25.7 * 10^(-3), nu= 15.3 * 10^(-6) , taxa="lizard_surface")

#Also illustrate estimations using a spherical approximation and a simplified version of the approximation.
heat_transfer_coefficient_approximation(V=V,D=svl,K= 25.7 * 10^(-3), nu= 15.3 * 10^(-6), taxa="lizard")
heat_transfer_coefficient_simple(V=0.5,D=0.05)

@@

We estimate convective heat exchange between the animal and surrounding air using ithe following relationship:
$$Q_{conv}= ef \cdot H_L(A\cdot \mbox{proportion})(T_a-T_b),$$
where an enhancement factor, $ef$, multiplier can be incorporated to account for increases in heat exchange resulting from air turbulance in field conditions

The function is available in R assuming that 2/3 of the lizard's surface area is exchanging heat through convection. We note we currently estimate no convection since we are assuming $Tb=Ta$ for illustration purposes:
```{r}
Qconv= rep(NA, 24)
for(hr in 1:24) Qconv[hr]= Qconvection(T_a= Ta[hr]+273.15,T_b= Ta[hr]+273.15,H=H_L,A=A, proportion=0.67, ef=1.3)
```

We estimate conductance between the lizard and surface assuming conductance through the animal tissue is the rate limiting step as follows:
$$Q_{cond}= A \cdot \mbox{proportion} \cdot K(T_g-T_b)/d, $$
We conduct the estimate in R assuming that 1/3 of the lizard surface is in contact with the ground and a skin thinkness of $10^{-6}$:
```{r}
Qcond= rep(NA, 24)
for(hr in 1:24) Qcond[hr]=Qconduction_animal(T_g= Ts[hr]+273.15,T_b=Ta[hr]+273.15,d=10^-6,K=K,A=A, proportion=0.33)
```

We assume, as in generally done for lizards, that heat exchange associated with metabolism and evaporation is negligible:
```{r}
Qmet= 0
Qevap= 0
```

The full heat budget can be calculated as follows: 
```{r}
Qnet_Gates(Qabs=Qabs, Qemit=Qemit, Qconv=Qconv, Qcond=Qcond, Qmet=Qmet, Qevap=Qevap)
```
We constructed the energy balance assuming $T_b=T_a$ for illustration. We now use a function based on the energy balance above to estimate body temperature given the environmental conditions:

```{r}
Te= rep(NA, 24)
for(hr in 1:24) Te[hr]= Tb_Gates(A=A, D=svl, psa_dir=psa[hr], psa_ref=1-psa[hr], psa_air=0.67, psa_g=0.33, T_g=Ts[hr]+273.15, T_a=Ta[hr]+273.15, Qabs=Qabs[hr], epsilon=epsilon_s, H_L=H_L, ef=1.3, K=K)
```

We also implement a similar but simplified energy balance. The energy balance omits conduction with the ground and thus underestimates temperatures, but we implement it for comparison purposes:
```{r}
Te2= rep(NA, 24)
for(hr in 1:24) Te2[hr]= Tb_CampbellNorman(T_a=Ta[hr]+273.15, S=Qabs[hr]/svl, epsilon=epsilon_s, c_p=29.3, D=svl, V=V)
#CHECK
```

We additionally estimate $T_e$ using the specialized function for lizards:
```{r}
Te3= rep(NA, 24)
for(hr in 1:24) Te3[hr]= Tb_lizard(T_a=Ta[hr], T_g=Ts[hr], u=V, svl=svl*1000, m=mass, psi=psi_deg[hr], rho_S=rho, elev=elev, doy=doy, sun=TRUE, surface=TRUE, alpha_S=a, alpha_L=0.965, epsilon_s=epsilon_s, F_d=0.8, F_r=0.5, F_a=0.5, F_g=0.5)

Te4= rep(NA, 24)
for(hr in 1:24) Te4[hr]= Tb_Fei(T_a=Ta[hr]+273.15, T_g=Ts[hr]+273.15, H=Qabs[hr], lw=Qabs[hr], shade=0, m=mass, Acondfact =0.1, Agradfact =0.3)
#CHECK

### Plot
We next plot the diurnal trends in environmental data and the operative data.

```{r}
par(mar=c(5, 4, 4, 8) + 0.1)
plot(1:24, (Te-273.15), type="l", xlab="Hour", ylab="Temperature (°C)", col="blue") #Gates
points(1:24, Te2-273.15, type="l", col="blue", lty="dotted") #Campbell
points(1:24, Te3, type="l", col="blue", lty="dashed") #Buckley 2008
#points(1:24, Te4, type="l", col="red", lty="solid") #Fei et al.
points(1:24, Ta, type="l", col="orange")
points(1:24, Ts, type="l", col="purple")

#add additional axis with radiation
par(new = T)
plot(1:24, Sdir, pch=16, axes=F, xlab=NA, ylab=NA, type="l", col="green")
axis(side = 4)
mtext(side = 4, line = 3, 'Radiation')
legend("topleft",
       legend=c("Ta", "Ts", "Te", "Radiation"),
       lty=1, pch=NA, col=c("orange", "purple", "blue", "green"))

## Summary
New environmental data and tools make biophysical modeling more accessible. Expand use since we realize.

>>>>

## Background
Many responses of organisms and ecosystems to climate change are inconsistent with current predictive models [@maguire2015modeling]. Some of these predictive shortcomings stem from omitting spatial and temporal environmental variation and how it interacts with organismal phenotypes (e.g., size, coloration) [@nadeau2017coarse; @potter2013microclimatic]. Many analyses assume that organismal body temperatures are equal to air temperate at weather station height (usually 2m). However, air temperatures near the ground where most organisms reside can be considerably warmer and absorption of solar radiation can raise body temperatures well above air temperatures, resulting in potential discrepancies of tens of degrees C [@kearney2009potential; @sinclair2016can; @sunday2014thermal].  Acute thermal stress events are often more relevant to the physiology and energy balance of organisms than daily, monthly, or annually aggregated environmental conditions. Yet, daily and seasonal environmental variation is often neglected [@dillon2016introduction; @dillon2016life]. We aim to help remedy these shortcomings by introducing the TrenchR package as a component of the TrEnCh project (https://trenchproject.github.io), which builds computational and visualization tools to Translate Environmental Change into organismal responses.

The mathematical tools for leveraging environmental data to predict organismal conditions have long lingered in books [@campbell2000introduction; @gates1980biophysical] and articles [@porter1969thermodynamic; @porter1983biophysical], but adoption of the tools has not kept pace with research on climate change responses. Adoptions of biophysical approaches are expanding with recognition of the importance of spatial and temporal environmental variation to determining climate change responses [@buckley2018leveraging; @kearney2017nichemapr].  Increasing dissemination of computational tools is facilitating adoption. The TrenchR package is intended to complement the NicheMapR package, which includes sophisticated microclimate [@kearney2017nichemapr] and biophysical [@kearney2020nichemapr] models. Many of the approaches are similar between the packages, but TrenchR focuses on simple functions to aid the understanding and transparency of biophysical approaches.  The TrenchR functions allow for modular use and easy adaptation to a variety of organismal and research questions.  Some energy budget calculation components of biophysical models are also available in the ThermImage package [@tattersall2019thermimage]. Although many of the TrenchR functions are general, some components of TrenchR are best suited for ectothermic animals. We have minimized inclusion of plant specific biophysical models since they are the focus of several R packages [tealeaves: [@muir2019tealeaves]; plantecophys [@duursma2015plantecophys]].

@@

## Examples
We illustrate the use of the TrenchR package by estimating an energy budget for a *Sceloporus* lizard on June 1, 2018, in Santa Fe, New Mexico, USA (35.69, -105.944, elevation: 2121m). The example is also incorporated in the *Using energy balances to estimate body temperatures* tutorial. We first estimate the energy budget with component functions and we then use an integrated biophysical model to estimate operative environmental temperatures. We start by generating environmental inputs. See Appendix S1 for another example examining a time series of microcliamte data and applying and testing the biophysical models.

@@

Let's assume the lizard is in an unshaded location where a weather station at standard height (2m) indicates that the daily air temperature varies from a minimum of 10°C to a maximum of 25°C and the wind speed averages 1 m/s. The soil surface temperature varies from a minimum of 15°C to a maximum of 30°C. We assume that atmospheric transmissivity $\tau=0.7$ and albedo $\rho=0.6$

@@

Tb_CampbellNorman=function(T_a, T_g, S, alpha_L=0.96, epsilon=0.96, c_p=29.3, D, V){

  stopifnot(T_a>200, T_a<400, epsilon>=0.5, epsilon<=1, c_p>=0, D>0, V>=0)

  #Stefan-Boltzmann constant
  sigma= 5.673*10^(-8) #W m^(-2) K^(-4)

  #solar and thermal radiation absorbed
  L_a=sigma*T_a^4  # (10.7) long wave flux densities from atmosphere 
  L_g=sigma*T_g^4  # (10.7) long wave flux densities from ground
  F_a=0.5; F_g=0.5 #proportion of organism exposure to air and ground, respectively
  R_abs= S+alpha_L*(F_a*L_a+F_g*L_g) # (11.14) Absorbed radiation

  #thermal radiation emitted
  Qemit= epsilon*sigma*T_a^4

  #conductance
  g_Ha=1.4*0.135*sqrt(V/D) # boundary conductance, factor of 1.4 to account for increased convection (Mitchell 1976), assumes forced conduction
  g_r= 4*epsilon*sigma*T_a^3/c_p # (12.7) radiative conductance

  # operative environmental temperature
  T_e=T_a+(R_abs-Qemit)/(c_p*(g_r+g_Ha))

  return(T_e) 
}
```

```{r}
#set up input data as variables
lat=35.69; lon= -105.944; elev=2121 #latitude and longitude (degrees) and elevation (m)
Tmin=10; Tmax=25; Tmin_s=15; Tmax_s=30 #minimum and maximum of air and soil temperatures, respectively (C)
V=1 #wind speed (m/s)
#assumptions
tau=0.7; rho=0.6 #atmospheric transmissivity and albedo
Tb0=25 #initial assumption of body temperature (C)

@@

Although measured solar radiation is preferable if available, we can estimate hourly solar radiation by discounting incoming solar radiation as it moves through the atmosphere as follows:
```{r}
psi_deg= sapply(1:24, FUN=zenith_angle, doy=doy, lat=lat, lon=lon) #estimate zenith angle (degrees)
psi_rad= degree_to_radian(psi_deg) #convert to radians
Srad= sapply(psi_rad, FUN=estimate_radiation, doy=doy, tau=tau, elev=elev, rho=rho) #estimate radiation
#Separate into direct, diffuse, and reflected solar radiation
Sdir= Srad[1,]; Sdif= Srad[2,]; Sref= Srad[3,]
```

We can then calculate hourly air and soil temperatures based on daily minimum and maximum temperatures. We select the sine-exponential model for air temperature and the sine model for surface temperature:
```{r}
Ta= sapply(1:24, diurnal_temp_variation_sineexp, T_max=Tmax, T_min=Tmin, t_r=tr, t_s=ts, alpha=2.59, beta= 1.55, gamma=2.2) #air temperature (C)
Ts= sapply(1:24, diurnal_temp_variation_sine, T_max=Tmax_s, T_min=Tmin_s) #soil temperature (C)
```

We use microclimate models to scale air temperature and wind speed from weather station height (2m) to lizard height (0.02m). We assume a surface roughness of $z_0=0.2$, which corresponds to bare sand and determines the turbulence of airflow. We implement neutral air temperature and wind speed profiles as well as an air temperature profile forced by wind speed.

```{r}
#scale air temperature to lizard height
#z is reference height, z is organism height, z0 is surface roughness
Ta_liz= air_temp_profile_neutral(T_r=Ta, zr=2, z0=0.2, z=0.02, T_s=Ts)
Ta_liz2= air_temp_profile(T_r=Ta, u_r=V, zr=2, z0=0.2, z=0.02, T_s=Ts)

#scale wind speed to lizard height
V_liz= wind_speed_profile_neutral(u_r=V, zr=2, z0=0.2,z=0.02)
```

### Energy balance
We will be solving the following energy balance to estimate $T_e$:
$$Q_{net} = Q_{abs} - Q_{emit} - Q_{conv} - Q_{cond} - Q_{met} - Q_{evap},$$
where $Q_{net}$ is the net energy exchange with the environment (W), $Q_{abs}$ is the solar radiation absorbed (W), $Q_{emit}$ is the net thermal radiation emitted (W), $Q_{conv}$ is energy exchange due to convection (W), $Q_{cond}$ is energy exchange due to conduction (W), $Q_{met}$ is the energy emitted due to metabolism (W), and $Q_{evap}$ is the energy emitted due to evaporative water loss (W). We will estimate each term on the right side of the equation in turn. Estimating $Q_{abs}$ requires the surface area exposed to radiation and the solar absorptivity of the animal surface ($a$ proportion). We model a 10g Sceloporus lizard with solar absorptivity $a=0.9$ [@gates1980biophysical]. We will initially assume $T_b=T_a$ to illustrate the calculations before solving for $T_b$ givne the environmental conditions.

```{r}
mass=10 #mass (g)
a=0.9 #solar absorptivity (%)

@@

We calculate the hourly solar radiation absorbed (W) as follows:
```{r}
Qabs= rep(NA, 24)
for(hr in 1:24) Qabs[hr]= Qradiation_absorbed(a=a, A=A, psa_dir=psa[hr], psa_ref=1-psa[hr], S_dir=Sdir[hr], S_dif=Sdif[hr], a_s=rho)
```

We estimate thermal radiation $Q_{emit}$ (W) for the lizard outdoors. We assume the surface emissivity of lizards, $epsilon_s=0.965$ [@barlett1967energy].

@@

Qemit= rep(NA, 24)
for(hr in 1:24) Qemit[hr]=Qemitted_thermal_radiation(epsilon=epsilon_s, A=A, psa_dir=psa[hr], psa_ref=1-psa[hr], T_b=Ta_liz[hr]+273.15, T_g=Ts[hr]+273.15, T_a=Ta_liz[hr]+273.15, enclosed=FALSE)
```

We next estimate convection $Q_{conv}$ (W) and conduction $Q_{cond}$ (W). We will estimate the lizard's heat transfer coefficient, $H_L$ ($Wm^{-2}K^{-1}$) using an empirical relationship for lizards (`heat_transfer_coefficient()`), but also illustrate a function estimating $H_L$ using a spherical approximation (`heat_transfer_coefficient_approximation()`) and a simplified approximation (`heat_transfer_coefficient_simple()`).

```{r}
#Use DRYAIR from NicheMapR to estimate the thermal conductivity of air and kinematic viscosity
ap= airpressure_from_elev(elev)*1000 #estimate Barometric pressure (pascal)

DRYAIRout= DRYAIR(db=Ta, bp=ap, alt=elev)
K= DRYAIRout$thcond #thermal conductivity (Wm^-2K^-1) 
nu= DRYAIRout$viskin #kinematic viscosity (m2 s-1)

svl=0.006 #approximate snout vent length (m) for Sceloporus

#We will use the average of K and nu across the day for simplicity and since there is not a substantial diurnal variation.
K= mean(K)
nu=mean(nu)

#Estimate the heat transfer coefficient using an empirical relationship for lizards
H_L=heat_transfer_coefficient(V=V_liz,D=svl,K= 25.7 * 10^(-3), nu= 15.3 * 10^(-6) , taxa="lizard_surface")

#Also illustrate estimations using a spherical approximation and a simplified version of the approximation.
H_L2= heat_transfer_coefficient_approximation(V=V_liz,D=svl,K= 25.7 * 10^(-3), nu= 15.3 * 10^(-6), taxa="lizard")
H_L3= heat_transfer_coefficient_simple(V=V_liz,D=svl)
```

We estimate convective heat exchange between the animal and surrounding air using the following relationship:
$$Q_{conv}= ef \cdot H_L(A\cdot \mbox{proportion})(T_a-T_b),$$
where an enhancement factor, $ef$, multiplier can be incorporated to account for increases in heat exchange resulting from air turbulence in field conditions. We implement the function in R assuming that 2/3 of the lizard's surface area is exchanging heat through convection. We note we currently estimate no convection since we are assuming $Tb=Ta$ for illustration purposes:
```{r}
Qconv= rep(NA, 24)
for(hr in 1:24) Qconv[hr]= Qconvection(T_a= Ta_liz[hr]+273.15, T_b= Ta_liz[hr]+273.15, H=H_L, A=A, proportion=0.67, ef=1.3)
```

We estimate conductive heat flow (W) from the lizard to the surface assuming conductance through the animal tissue is the rate limiting step as follows:
$$Q_{cond}= A \cdot \mbox{proportion} \cdot K(T_g-T_b)/d. $$
We implement the estimate in R assuming that 1/3 of the lizard surface is in contact with the ground and a skin thickness $d$ of $10^{-6}$:
```{r}
Qcond= rep(NA, 24)
for(hr in 1:24) Qcond[hr]=Qconduction_animal(T_g= Ts[hr]+273.15, T_b=Ta_liz[hr]+273.15, d=10^-6, K=K, A=A, proportion=0.33)
```

We assume, as in generally done for lizards, that heat exchange associated with metabolism and evaporation is negligible:
```{r}
Qmet= 0; Qevap= 0
```

The full heat budget can be calculated as follows [@gates1980biophysical]: 
```{r}
Qnet= Qnet_Gates(Qabs=Qabs, Qemit=Qemit, Qconv=Qconv, Qcond=Qcond, Qmet=Qmet, Qevap=Qevap)
```
We constructed the energy balance assuming $T_b=T_a$ for illustration. Thus, the energy balance is dominated by heat gain from contact with the warmer surface. We now use a function based on the energy balance above to estimate body temperature given the environmental conditions:

```{r}
Te= rep(NA, 24)
for(hr in 1:24) Te[hr]= Tb_Gates(A=A, D=svl, psa_dir=psa[hr], psa_ref=1-psa[hr], psa_air=0.67, psa_g=0.33, T_g=Ts[hr]+273.15, T_a=Ta_liz[hr]+273.15, Qabs=Qabs[hr], epsilon=epsilon_s, H_L=H_L, ef=1.3, K=K)
```

We also implement a similar but simplified energy balance [@campbell2000introduction]. The energy balance omits conduction with the ground:
```{r}
Te2= rep(NA, 24)

for(hr in 1:24) Te2[hr]= Tb_CampbellNorman(T_a=Ta_liz[hr]+273.15, T_g=Ts[hr]+273.15, S=Qabs[hr]/A, alpha_L=0.96, epsilon=epsilon_s, c_p=29.3, D=svl, V=V_liz)
#S is solar radiation flux (W m^-2), so we divide by surface area, A
```

We additionally estimate $T_e$ using a specialized function for lizards [@buckley2008linking]:
```{r}
Te3= rep(NA, 24)
for(hr in 1:24) Te3[hr]= Tb_lizard(T_a=Ta_liz[hr], T_g=Ts[hr], u=V_liz, svl=svl*1000, m=mass, psi=psi_deg[hr], rho_S=rho, elev=elev, doy=doy, sun=TRUE, surface=TRUE, alpha_S=a, alpha_L=0.965, epsilon_s=epsilon_s, F_d=0.8, F_r=0.5, F_a=0.5, F_g=0.5)
```

### Plot
We next plot the diurnal trends in environmental data and predicted body temperatures.

```{r}
par(mar=c(5, 4, 4, 8) + 0.1)
plot(1:24, (Te-273.15), type="l", xlab="Hour", ylab="Temperature (°C)", col="#5DC863FF", ylim=c(10,60)) #Gates
points(1:24, Te2-273.15, type="l", col="#5DC863FF", lty="dotted") #Campbell and Norman
points(1:24, Te3, type="l", col="#5DC863FF", lty="dashed") #Lizard, Buckley 2008
points(1:24, Ta, type="l", col="#440154FF")
points(1:24, Ta_liz, type="l", col="#3B528BFF")
points(1:24, Ts, type="l", col="#21908CFF")

#add additional axis with radiation
par(new = T)
plot(1:24, Sdir, pch=16, axes=F, xlab=NA, ylab=NA, type="l", col="#FDE725FF")
axis(side = 4)
mtext(side = 4, line = 3, 'Radiation (W/m^2)')
legend("topleft",
       legend=c("Ta 2m", "Ta 0.02m", "Ts", "Te", "Radiation"),
       lty=1, pch=NA, col=c("#440154FF", "#3B528BFF", "#21908CFF", "#5DC863FF", "#FDE725FF"))

@@

*Figure 1*. Predicted lizard body temperatures (Te) as a function of air temperatures (Ta) at 2m and lizard height (0.02m), surface (Ts) temperatures (°C), and solar radiation ($W/m^2$). We depict body temperatures using two general energy budgets [solid: Tb_Gates(); dotted: Tb_CampbellNorman()] and a lizard specific biophysical model [dashed: Tb_lizard()].

The microclimate models indicate that air temperatures at lizard height are similar to surface temperatures (Figure 1). The biophysical models indicate that solar radiation will elevate lizard body temperatures far above air temperatures and that the lizard will face thermal stress if it is unable to seek shade (Figure 1). The three biophysical models predict different body temperatures during peak period of solar radiation because they model interactions with the ground differently and users are encouraged to review the details of each biophysical model before selection.

## Summary
TrenchR is intended to promote understanding of how organisms interact with their environment and the consequences for physiology, energetics, behavior, and demography. Our example implementation highlights the importance of considering organismal body temperatures, rather than air temperatures, when examining thermal stress and other responses to environmental variability and change. TrenchR currently focuses on heat balances but may be expanded to include water balances. Simple functions can be combined as needed to produce comprehensive and transparent models for biophysical ecology and evolution. The resultant models are likely to be sufficiently detailed for many applications, but users are referred to NicheMapR for more detailed biophysical models [@kearney2017nichemapr; @kearney2020nichemapr]. Making classic biophysical ecology techniques more accessible will allow researchers to take advantage of rapidly accumulating data on environmental conditions and organismal traits to understand and predict ecological and evolutionary responses to variable and changing environments.
